Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis

### 2022-04-01 Manuscript/SS_GAM_Supporting_Information.tex
Add Supporting information to workflow

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{Spike-and-Slab Generalized Additive Models and Scalable
Algorithms for High-Dimensional Data}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supporting Infromation}
\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
Nengjun Yi}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data},
  pdfauthor={Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long, Nengjun Yi},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

Given that
\(\gamma_{j}^*| \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)\)
where \(\gamma_{j} | \theta_j \sim Bin(1, \theta_j)\), we can derive the
the marginal distribution of \(\gamma_{j}^*\) with the following
manipulation. \begin{align*}
& \text{Pr}(\gamma_j^*= 1 | \theta_j)  = \text{Pr}(\gamma_j^*= 1, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 1 , \gamma_j = 0| \theta_j)\\
= & \text{Pr}(\gamma_j^*= 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^*\text{ and }\gamma \text{.]}\\
= & \text{Pr}(\gamma_j^*= 1| \gamma_j = 1,  \theta_j)\text{Pr}(\gamma_j = 1| \theta_j)\\
= & \theta_j^2\\
& \text{Pr}(\gamma_j^*= 0 | \theta_j)  = \text{Pr}(\gamma_j^*= 0, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 0 , \gamma_j = 0| \theta_j)\\
= & \text{Pr}(\gamma_j^*= 0, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 0, \gamma_j = 0 | \theta_j)\\
= & \text{Pr}(\gamma_j^*= 0| \gamma_j = 1,  \theta_j)\text{Pr}(\gamma_j = 1| \theta_j) + \text{Pr}(\gamma_j^*= 0| \gamma_j = 0,  \theta_j)\text{Pr}(\gamma_j = 0| \theta_j)\\
= & (1-\theta_j)\theta_j + 1(1-\theta_j) = 1-\theta_j^2
\end{align*}

@@

Similar to the EM-CD algorithm, the EM-Iterative Weighted Least Square
(EM-IWLS) algorithm is an EM-based algorithm where the iterative
weighted least squares algorithm is used to find the estimate of
\(\boldsymbol{\beta}, \phi\) that maximizes \(E(Q_1)\). The iterative
weighted least squares algorithm was originally proposed to fit the
classical generalized linear models, and generalized to fit some
Bayesian hierarchical models.{[}@Gelman2013{]} Yi and Ma {[}@Yi2012{]}
extended the algorithm to fit Bayesian hierarchical models for
high-dimensional data analysis. Specifically, the authors formulated
Student's t-distribution and double exponential distribution as
hierarchical normal distributions such that generalized linear models
with shrinkage priors can be easily fitted. In this work, we further
extend the EM-IWLS paradigm to fit the proposed BHAM method. COmpare to
the EM-CD algorithm, EM-IWLS estimate the variance-covariance matrix of
the coefficients, providing an opportunity to derive the uncertainty
quantification of smoothing functions. We will defer this discussion to
the future work, due to the delicacy of the topic. However, the EM-IWLS
is not as computationally efficient as EM\_CD, particularly in
high-dimensional settings.

@@

  \begin{enumerate}
  \def\labelenumii{\alph{enumii})}
  \tightlist
  \item
    Based on the current value of \(\beta\), calculate the pseudo-data
    \(z_i^{(t)}\) and the pseudo-weights \(w_i^{(t)}\)
  \item
    Update \(\boldsymbol{\beta}^{(t)}\) by runing the augmented weighted
    least squared
  \item
    If \(\phi\) is present, update \(\phi\)
  \end{enumerate}
\end{enumerate}

@@

\subsection{Supplementary Information 4: Variable Selection Performance of Simulations}
\input{Tabs/sim_binom_var_slct_tab.tex}
\input{Tabs/sim_lnr_gaus_var_slct_tab.tex}
\input{Tabs/sim_lnr_binom_var_slct_tab.tex}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{Spike-and-Slab LASSO Generalized Additive Models and Scalable
Algorithms for High-Dimensional Data Analysis}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supporting Information}
\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
Nengjun Yi}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis},
  pdfauthor={Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long, Nengjun Yi},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

Given that
\(\gamma_{j}^*| \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)\)
where \(\gamma_{j} | \theta_j \sim Bin(1, \theta_j)\), we can derive the
marginal distribution of \(\gamma_{j}^*\) with the following
manipulation. \begin{align*}
& \text{Pr}(\gamma_j^*= 1 | \theta_j)  = \text{Pr}(\gamma_j^*= 1, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 1 , \gamma_j = 0| \theta_j)\\
= & \text{Pr}(\gamma_j^*= 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^*\text{ and }\gamma \text{.]}\\
= & \text{Pr}(\gamma_j^*= 1| \gamma_j = 1,  \theta_j)\text{Pr}(\gamma_j = 1| \theta_j)\\
= & \theta_j^2\\
& \text{Pr}(\gamma_j^*= 0 | \theta_j)  = \text{Pr}(\gamma_j^*= 0, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 0 , \gamma_j = 0| \theta_j)\\
= & \text{Pr}(\gamma_j^*= 0, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 0, \gamma_j = 0 | \theta_j)\\
= & \text{Pr}(\gamma_j^*= 0| \gamma_j = 1,  \theta_j)\text{Pr}(\gamma_j = 1| \theta_j) + \text{Pr}(\gamma_j^*= 0| \gamma_j = 0,  \theta_j)\text{Pr}(\gamma_j = 0| \theta_j)\\
= & (1-\theta_j)\theta_j + 1(1-\theta_j) = 1-\theta_j^2
\end{align*}

@@

Similar to the EM-CD algorithm, the EM-Iterative Weighted Least Square
(EM-IWLS) algorithm is an EM-based algorithm where the iterative
weighted least squares algorithm is used to find the estimate of
\(\boldsymbol{\beta}, \phi\) that maximizes \(E(Q_1)\). The iterative
weighted least squares algorithm was originally proposed to fit the
classical generalized linear models, and generalized to fit some
Bayesian hierarchical models.(Gelman et al. 2013) Yi and Ma (Yi and Ma
2012) extended the algorithm to fit Bayesian hierarchical models for
high-dimensional data analysis. Specifically, the authors formulated
Student's t-distribution and double exponential distribution as
hierarchical normal distributions such that generalized linear models
with shrinkage priors can be easily fitted. In this work, we further
extend the EM-IWLS paradigm to fit the proposed BHAM method. Compare to
the EM-CD algorithm, EM-IWLS estimates the variance-covariance matrix of
the coefficients, providing an opportunity to derive the uncertainty
quantification of smoothing functions. We will defer this discussion to
future work, due to the delicacy of the topic. However, the EM-IWLS is
not as computationally efficient as EM-CD, particularly in
high-dimensional settings.

@@

  \begin{enumerate}
  \def\labelenumii{\alph{enumii})}
  \tightlist
  \item
    Based on the current value of \(\beta\), calculate the pseudo-data
    \(z_i^{(t)}\) and the pseudo-weights \(w_i^{(t)}\)
  \item
    Update \(\boldsymbol{\beta}^{(t)}\) by running the augmented
    weighted least squared
  \item
    If \(\phi\) is present, update \(\phi\)
  \end{enumerate}
\end{enumerate}

@@

\subsection{Supplementary Information 5: Variable Selection Performance of Simulations}
\input{Tabs/sim_binom_var_slct_tab.tex}
\input{Tabs/sim_lnr_gaus_var_slct_tab.tex}
\input{Tabs/sim_lnr_binom_var_slct_tab.tex}

\clearpage
\subsection{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Gelman2013}{}}%
Gelman, A, JB Carlin, HS Stern, DB Dunson, A Vehtari, and BD Rubin.
2013. {``Bayesian Data Analysis. 3rd Editio.''}

\leavevmode\vadjust pre{\hypertarget{ref-Yi2012}{}}%
Yi, Nengjun, and Shuangge Ma. 2012. {``{Hierarchical Shrinkage Priors
and Model Fitting for High-dimensional Generalized Linear Models}.''}
\emph{Statistical Applications in Genetics and Molecular Biology} 11
(6). \url{https://doi.org/10.1515/1544-6115.1803}.

\end{CSLReferences}
